---
title: "PredictiveAssignment5"
author: "Christine Iyer"
date: "April 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


```{r}
library(forecast)
library(knitr)
library(dplyr)
library(pander)
library(Hmisc)
```
###1. Canadian Manufacturing

1. Analysis of Canadian Manufacturing Workers Work-Hours: The time series plot in Figure 7.9 describes the average annual number of weekly hours spent by Canadian manufacturing workers. The data is available in CanadianWorkHours.xls.

```{r, include=FALSE}
setwd("/Users/Chris Iyer/Documents/Assignment5Predictive/")
canadian <- read.csv("CanadianWorkHours.csv", header = TRUE, stringsAsFactors = FALSE)
names(canadian)
names(canadian)[2] <- "Hours"
head(canadian)
#convert to a time series
canadian.ts <- ts(canadian$Hours, start=c(1966, 1), frequency=1)
```


```{r, include=FALSE}
#partition the data
rangeCA <- range(canadian.ts)
#rangeCA
#35 years. 
length(canadian.ts)
validLength <- 5
trainLength <- length(canadian.ts) - validLength
windowTrain <- window(canadian.ts, start = c(1966,1), end = c(1966, trainLength))
windowValid <- window(canadian.ts, start = c(1966, trainLength + 1), end =  c(1966, trainLength + validLength))
```


```{r, include=FALSE}
linearTrain <- tslm(canadian.ts ~ trend)
summary(linearTrain)
```


**This shows the linear trend plotted on the generic plot**

```{r, echo=FALSE}
yrange <- range(canadian.ts)
#plot ts and linear trend
plot(c(1966, 2000), yrange, type="n", xlab="Year",  ylab = "Average Weekly Hours Worked",  main = "Average Weekly Hours Worked in Canadian Manufacturing", bty="l", xaxt="n", yaxt="n")
# Add the x-axis
axis(1, at=seq(1966,2000,1), labels=format(seq(1966,2000,1)))
# Add the y-axis
axis(2, at=seq(30,40, 0.5), labels=format(seq(30,40, 0.5)), las=2)
# Add the time series canadian
lines(canadian.ts, bty="l")
```



**(a) If we computed the autocorrelation of this series, would the lag-1 autocorrelation exhibit negative, positive, or no autocorrelation? How can you see this from the plot?** 

The time series plot of the averge weekly hours worked in Canadian manufacturing largely shows a downlward trend. If we computed a lag-1 autocorrelation, we would see "stickiness", consecutive values that move in same direction, be it up or down. This is a positive autocorrelation. When fitting a model to this data series, we would want to look for one that either captures trend or we would detrend the data first. 



**(b) Compute the autocorrelation and produce an ACF plot. Verify your answer to the previous question.**

It exhibits a strong positive autocorrelation at a lag 1. This is apparent from the line at period 1 on the x axis that crosses the significance threshold and goes nearly to 1 on the y axis. 

```{r}
canadaAcf <- Acf(canadian.ts, lag.max = 1, main = "Auto Correlation \nCanadian Work Hours")
canadaAcf
```


###2. Wal-Mart Stock

2. Forecasting Wal-Mart Stock: Figure 7.10 shows a time plot of Wal-Mart daily closing prices between February 2001 and February 2002. The data is available at finance.yahoo.com and in WalMartStock.xls. 21 The ACF plots of these daily closing prices and its lag-1 differenced series are in Figure 7.11. Table 7.4 shows the output from fitting an AR(1) model to the series of closing prices and to the series of differences. Use all the information to answer the following questions.

```{r, echo=FALSE}
setwd("/Users/Chris Iyer/Documents/Assignment5Predictive/")
walMart <- read.csv("WalMartStock1.csv", header = TRUE, stringsAsFactors = FALSE)
walMart.ts <- ts(walMart$Close, start = c(2001, 1), frequency = 248)

walMartLinearTrain <- tslm(walMart.ts ~ trend + I(trend^2))
#summary(walMartLinearTrain)

yrange <- range(walMart.ts)
# Set up the plot
plot(c(2001, 2002), yrange, type="n", xlab="Year 2001",  ylab = "Daily Closing Prices", main = "Daily Closing Prices \nWalMart", bty="l", xaxt="n", yaxt="n")

# Add the time series canadian
lines(walMart.ts, bty="n", col = "black")
#lines(walMartLinearTrain$fitted.values , col = "blue")

# Add the x-axis

axis(1, at=seq(2001,2001+11/12,1/12), labels=c("Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Jan"))
# Add the y-axis
axis(2, at=seq(40,65, 2), labels=format(seq(40,65, 2)), las=2)
#legend(2001,60, c("Actual", "Quad Trend"), lty=c(1, 1), col=c("black", "blue"), bty="n")
```


```{r}
Acf(walMart.ts, main = "ACF forWalMart Series", lag.max = 10)
Acf(diff(walMart.ts, lag=1),  main="ACF Plot for Differenced WalMart Series", lag.max = 10)
```


**(a) Create a time plot of the differenced series.**

```{r}
plot(diff(walMart.ts, lag = 1), main = "Differenced WalMart")

```


**(b) Which of the following is/are relevant for testing whether this stock is a random walk?**


* The autocorrelations of the closing price series **Relevant**

* The AR(1) slope coefficient for the closing price series **Relevant**
If the AR(1) slope coefficient $\beta_1 = 1$, the series is a random walk. 

* The AR(1) constant coefficient for the closing price series 

* The autocorrelations of the differenced series **Relevant**

* The AR(1) slope coefficient for the differenced series **Relevant**

* The AR(1) constant coefficient for the differenced series 

**(c) Recreate the AR(1) model output for the Close price series shown in the left of Table 7.4. Does the AR model indicate that this is a random walk? Explain how you reached your conclusion.**

To test whether this series is a random walk, the $\beta_1 = 1$. We can determine this by performing an hypothesis test where the $H_0: \beta_1 = 0$ and $H_1: \beta_1\ne 1$. The p value is calculated using the coefficient and standard error output from the AR(1) model.  At $\alpha = 0.01$, the AR(1) data indicates the Walmart closing price data series is a random walk because $p>\alpha$ using both the t distribution and the normal distribution. However, at $\alpha = 0.05$, we can reject the null hypothesis and conclude there is an underlying pattern and we can ultimately use a forecasting technique beyone the naive method. 

I've recreated the AR(1) model output. 


```{r}
walMartFit <- arima(walMart.ts, order = c(1,0,0))
walMartFit

```

**The p-value using the t distribution**

```{r}
#Calculate the two-tailed p-value from a t-distribution, p = 0.01896261
pander(2*pt(-abs((1 - walMartFit$coef["ar1"]) / 0.0187), df=length(walMart.ts)-1))
```

**The p-value using a normal distribution**

```{r}
#Now calculate it using the normal distribution, p = 0.01818593
pander(2*pnorm(-abs((1 - walMartFit$coef["ar1"]) / 0.0187)))
```

**(d) What are the implications of finding that a time series is a random walk? Choose the correct statement(s) below.**

* It is impossible to obtain useful forecasts of the series.

* The series is random. 

* The changes in the series from one period to the other are random. **This is a correct statement**

###3. Souvenir Sales

**3. Souvenir Sales: The file SouvenirSales.xls contains monthly sales for a souvenir shop at a beach resort town in Queensland, Australia, between 1995 and 2001.** 

**Back in 2001, the store wanted to use the data to forecast sales for the next 12 months (year 2002). They hired an analyst to generate forecasts. The analyst first partitioned the data into training and validation periods, with the validation set containing the last 12 months of data (year 2001). She then fit a regression model to sales, using the training period.**


```{r, echo=FALSE}
setwd("/Users/Chris Iyer/Documents/Assignment5Predictive/")
souvenirSales<- read.csv("SouvenirSales.csv")
souvenir.ts <- ts(souvenirSales$Sales, start = c(1995,1), frequency = 12)
validLength <- 12
trainLength <- length(souvenir.ts) - validLength
souvenirTrain <- window(souvenir.ts, end=c(1995, trainLength))
souvenirValid <- window(souvenir.ts, start=c(1995,trainLength+1))
```

**Log(souvenir) plot**

```{r}
plot(log(souvenir.ts), type = "n", xaxt = "n", yaxt = "n", xlab = "Year", ylab = "Souvenir Sales (thousands)", main = "Log Souvenir Sales")
lines(log(souvenir.ts), col = "grey48", lwd = 2, lty = 1, pch = 18, type = "o")
# Add the x-axis
axis(1, at=seq(1995,2002,1), labels=format(seq(1995,2002,1)))
# Add the y-axis
axis(2, at = seq(7, 11, 1), labels = format(round(exp(seq(7, 11, 1)), digits = 0), las = 2))
```

**(a)	Run a regression model with log(Sales) as the output variable and with a linear trend and monthly predictors. Remember to fit only the training period. Use this model to forecast the sales in February 2002.** 

**Log Regression trend and seasonality**

```{r}
LogFC <- tslm(log(souvenirTrain) ~ trend + season)
pander(summary(LogFC))
```
 
**Forecast using a regression model with log(Sales) as the output variable and with a linear trend and monthly predictors**  

```{r}
FCFC <- forecast(LogFC, h = validLength)
```

**(b) Create an ACF plot until lag-15 for the forecast errors. Now fit an AR model with lag-2 [ARIMA(2, 0, 0)] to the forecast errors.** 

```{r}
Acf(FCFC$residuals, lag.max = 15, main = "ACF plot of Forecast Errors \nLag = 15")
Arima2 <- Arima(LogFC$residuals, order = c(2,0,0))
Arima2
```

Running an AR(2) model shows there might be some trend which can be further modeled. To determine this we need to see if these AR(2) findings are statistically significant.

**Tests for significance**

```{r}
# Calculate the t statistics for each: coefficient / s.e.
# Rough rule is anything > 2 (or < -2) is statistically significant
#AR(1) is statistically significant 2.818482
Arima2$coef["ar1"] / 0.1090
#AR(2) is statistically significant at 3.346186
Arima2$coef["ar2"] / 0.1102
#both t statistics are statistically significant, so there is trend remaining which might be worth correcting the forecast for.
#estimate the p values ar1 p value < alpha, 0.004825136
pander(2*pnorm(-abs(Arima2$coef["ar1"]/0.1090)))
#p = 0.0008193151, p<alpha, 
pander(2*pnorm(-abs(Arima2$coef["ar2"]/0.1102)))
```

**The Acf of the AR(2)residuals**

```{r}
Acf(Arima2$residuals, main = "Acf of AR(2) Residuals")
```

**i.	Examining the ACF plot and the estimated coefficients of the AR(2) model (and their statistical significance), what can we learn about the regression forecasts?** 

Looking at the ACF plot and the estimated coefficients for the forecast errors, we see some trend that was missed in the forecast. Specifically at lags 1 and 2. This is reinforeced with the AR(2) model of the residuals; further modeling may take care of the remaining trend. We can perform hypothesis testing using the coefficients and standard errors to calculate t statistics. Above, I calculate the t-statistics for the ar1 and ar2. Both are statistically significant a because they are greater than 2. I further calculate the p values using the coefficients and standard errors. Using a normal distribution, both p values of ar1 and ar2 are smaller than an $\alpha = 0.01$. With a statistically significant AR2 model, we see there is some trend that can be further modeled to improve the Souvenir Sales forecast. The Acf of the AR(2) model shows that the remaining trend in the residuals can be modeled with the Arima method.  


**ii.	Use the autocorrelation information to compute a forecast for January 2002, using the regression model and the AR(2) model above.**



```{r}
FCArima <- forecast(Arima2, h = validLength)
ErrorError <- exp(FCFC$mean + FCArima$mean)
ErrorError[1]
```

**Plot the actual souvenir sales, the original forecast using a regression model with trend and seasonality using a log(souvenir sales)**


```{r}
plot(souvenirValid, bty = "l", xaxt = "n", yaxt = "n", xlab = "2001", ylab = "Souvenir Sales (thousands)", main = "Actual Sales vs Forecasted Sales \nSouvenir Sales", ylim = c(9000, 110000), lwd = 2)
axis(1, at=seq(2001,2001+11/12,1/12), labels= c("Jan ", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
axis(2, at = seq(0, 110000, 20000), labels = format(seq(0, 110, 20)), las = 2)
lines(exp(FCFC$mean), col = "blue", lwd = 2)
lines(ErrorError, col = "green", lwd = 2, lty = 2)
legend(2001, 90000, c("Actual Sales", "Forecasted Sales MAPE = 15.52", "Adjusted Forecasted Sales MAPE = 15.35"), lty = c(1,1,2), lwd = c(2,2,2), bty = "n", col = c("black", "blue", "green"))
```

**The predictive accuracy between the adjusted forecast and the original forecast shows a minor difference**

```{r}
accuracy(ErrorError, souvenirValid)
accuracy(exp(FCFC$mean), souvenirValid)
```


###4. Shipments of Household Appliances

**4. Shipments of Household Appliances: The file ApplianceShipments.xls contains the series of quarterly shipments (in millions of USD) of U.S. household appliances between 1985 and 1989. The series is plotted in Figure 7.13.**

```{r load ApplianceShipments1}
appliance <- read.csv("ApplianceShipments1.csv", header = TRUE, stringsAsFactors = FALSE)
appliance <- appliance %>% select(Quarter, Shipments)
#split Quarter column
col1Split <- strsplit(appliance$Quarter, "-")
#Make quarter column into proper columns in a dataframe
tempDF <- data.frame(matrix(unlist(col1Split), nrow = length(col1Split), byrow = TRUE), appliance$Shipments)
#reorder rows by year and then by quarter
tempDF <- tempDF[order(tempDF$X2, tempDF$X1),]
#merge quarter and year back together
m <- paste0(tempDF$X1, "-", tempDF$X2)
#put a new dataframe together
newDF <- data.frame(m, tempDF$appliance.Shipments)
#rename columns
names(newDF)[1] <- "Quarter"
names(newDF)[2] <- "Shipments"
#make this into a timeseries
applianceShipments.ts <- ts(newDF$Shipments, start = c(1984, 1), end = c(1989,4), frequency = 4)
plot(applianceShipments.ts, type = "o")
```


```{r}
library(Hmisc)
```

**(a) If we compute the autocorrelation of the series, which lag (> 0) is most likely to have the largest coefficient (in absolute value)?**

4

**(b) Create an ACF plot and compare it with your answer.**

```{r}
Acf(applianceShipments.ts, lag.max = 4)
Acf(diff(applianceShipments.ts, lag = 1))
```





```{r}
time <- time(applianceShipments.ts)
appValidLength <- 4
appTrainLength <- length(applianceShipments.ts) - appValidLength
#windows for analysis
appWindowTrain <- window(applianceShipments.ts, start = time[1],end = time[appTrainLength])
appWindowValid <- window(applianceShipments.ts, start = time[appTrainLength + 1], end = time[appTrainLength + appValidLength])
#applExpoSeason <- tslm(applianceShipments.ts ~ trend + season, lambda = 0)
applExpoSeason <- tslm(appWindowTrain ~ trend + season, lambda = 0)
#salesForecast <- forecast(applExpoSeason, h = 4)
summary(applExpoSeason)
#ggseasonplot(applianceShipments.ts)
```

**Plot with exp trend and seasonality**

```{r, echo=TRUE}
yrange <- range(appWindowTrain)
# Set up the plot
plot(c(1985, 1989), yrange, type="n", xlab="Year",  ylab="Appliance Sales", bty="l", xaxt="n", yaxt="n",  main = "Exponential Trend  with Seasonality Model \nfit to Appliance Store Sales ")
# Add the time series canadian training set
lines(appWindowTrain, bty="l", type = "l")
#add exponential fit
lines(applExpoSeason$fitted.values, bty = "l", col = "blue")
#lines(salesForecast$mean, col = "blue", lty = 2)
# Add the x-axis
axis(1, at=seq(1985,1989,1), labels=format(seq(1985,1989,1)), minor.tick(nx = 4))
# Add the y-axis
axis(2, at=seq(3000, 5000, 200), labels=format(seq(3000, 5000, 200)), las=2)
legend(1, 900, c("Appliance Store Sales", "Exp Trend with Seasonality Fit"
                   #, "Expo Forecast"
                   ), col = c("black", "blue"), lwd = c(1,1),  bty = "n", lty = c(1,1))
#accuracy(salesForecast, appWindowValid)
```


**Residual plot**


```{r}
plot(appWindowTrain - applExpoSeason$fitted.values, type="o", bty="l", xlab = "Year", ylab = "Residuals", main = "Appliance Sales Residuals in Training Period \nExponential Trend with Seasonality Fit")
abline(h = 0)
```




###Extras
```{r}

```{r}
salesLinearLog <- tslm(souvenirTrain ~ trend + season, lambda = 0)
summaryLinearLog <- summary(salesLinearLog)
#modelC summary
summaryLinearLog
forecastLinearLog <- forecast(salesLinearLog, h = validLength)
#AccC <- accuracy(FCC$mean, souvenirValid)
# The forecast
forecastLinearLog
#The accuracy of the forecast
#pander(AccC)
```
```

